Citation Analysis of Computer Science Publications

### 2019-12-11 Scientometrics/comp.tex
added cross-match figure

Progressing beyond 1974-79, and Salton and Bergmark's historical triad of theoretical foundations; hardware and computer systems; and software, the Computing Classification System (CCS)~\cite{acm_ref} published by the Association for Computing Machinery now consists of 13 top-level areas that reflect a more current view of the field. This classification also addresses relationships with other fields under the category of Applied Computing. However, an easy way to map scientific publications, especially interdisciplinary articles or those from fields proximal to the CCS classification does not exist.

@@

For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography published in the years 1996-2015, that had (i) a digital object identifier (DOI), and (ii) could also be matched to article identifiers in the Scopus bibliography. The process of selection and matching resulted in a dataset of 2,685,356 publications (Materials and Methods). Of these roughly 2.8 million publications, approximately 2.07 million were classified as COMP under subject area and 2.07 million similarly were assigned ASJC codes corresponding to Computer Science (major subject area), with the balance of 610,000 publications non-exclusively shared between 26 different major subject areas ranging from Engineering (330,048) to Dentistry (1). Thus, the sets of computer science publications in DBLP and Scopus are partially overlapping rather than DBLP being a perfect subset of Scopus. [the point being that we should justify our use of both journal and article level classifications].

\subsection{Clustering by Direct Citation} 
Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation with an argument being made for direct citation as the best approach to concentrate citation links~\cite{klavans_which_2017}. Accordingly, we first used direct citation links as the basis for cluster formation. A set of 20 clusters was generated using Graclus, a spectral clustering technique that has previously been applied to citation data~\cite{graclus_2007,subelj_clustering_2016}. 
\textcolor{blue}{
Considering the caveats associated with generating \emph{comp}, namely, the coverage limitations of the DBLP dataset (ii) cross-matching to Scopus (iii) journal based classification (iv) clustering, }
we constructed article level clusters of this computer science dataset at a sufficiently high level of aggregation to avoid cognitive challenge and cross-matched them to disciplinary classification by journal, the Scopus ASJC classification. To focus on relatively high signal, we only considered Scopus ASJC minor subject area categories that accounted for at least 15\% of the publications in each cluster.

@@

We also attempted to map these 20 clusters assembled through direct citation to the CCS classification by manually matching the top 25-50 most heavily cited publications in each cluster to corresponding categories in the CCS. This was feasible in some cases, clusters 0 and 1 mapped reasonably well to the top level categories Hardware and Computer Systems Organization but less so in the cases of other clusters. Further, the top cited papers often derived from biology and we were concerned that this small sample (top cited papers) were not representative of the cluster.

\subsection{Clustering by Co-citation.}

@@

@@

\begin{figure}[ht]
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.45]{scopus_dblp_heatmap3.pdf}
% figure caption is below the figure
\caption{Heat map for the clustering obtained by direct citation. The  y-axis (rows) correspond to topics, defined by Scopus characterizations, and the
x-axis (columns) represent the 20 different clusters.  
Each cluster is characterized by topics that label at least 10\% of the publications in the cluster.
 }
\label{fig:heatmap}       % Give a unique label
\end{figure}

@@

>>>>

Beyond Salton and Bergmark's historical triad of theoretical foundations; hardware and computer systems; and software, the Computing Classification System (CCS)~\cite{acm_ref} published by the Association for Computing Machinery now consists of 13 top-level areas that reflect a more current view of the field. This classification also addresses relationships with other fields under the category of Applied Computing. However, an easy way to map scientific publications, especially interdisciplinary articles or those from fields proximal to the CCS classification does not exist.

@@

For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography published in the years 1996-2015, that had (i) a digital object identifier (DOI), and (ii) could also be matched to article identifiers in the Scopus bibliography. The process of selection and matching resulted in a dataset of 2,685,356 publications (Materials and Methods). Of these roughly 2.8 million publications, approximately 2.07 million were classified as COMP under subject area and assigned ASJC codes corresponding to Computer Science (major subject area and 13 minos subject areas), with the balance of 610,000 publications non-exclusively shared between 26 different major subject areas ranging from Engineering (330,048) to Dentistry (1). The set of 2.07 million publications classified as COMP in Scopus spanned all 13 minor subject areas with Software 30.3\% being the largest component and Computer Science (miscellaneous) 0.9\% the smallest. The number of minor subject areas assigned to a publication averaged to 1.75 with 53\% of the publications being assigned to one minor subject area and 0.0003\% assigned to 7 different minor subject areas (Table 1).

Thus, the sets of computer science publications in DBLP and Scopus are partially overlapping rather than DBLP being a perfect subset of Scopus. [the point being that we should justify our use of both journal and article level classifications].

\subsection{Clustering by Direct Citation} [Need a sentence about why we're clustering].
Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation with an argument being made for direct citation as the best approach to concentrate citation links~\cite{klavans_which_2017}. Accordingly, we first used direct citation links as the basis for cluster formation. A set of 20 clusters was generated using Graclus, a spectral clustering technique that has previously been applied to citation data~\cite{graclus_2007,subelj_clustering_2016}. 
\textcolor{blue}{
Considering the caveats associated with generating \emph{comp}, namely, the coverage limitations of the DBLP dataset (ii) cross-matching to Scopus (iii) journal based classification (iv) clustering, }
we constructed article level clusters of this computer science dataset at a sufficiently high level of aggregation to avoid cognitive challenge and cross-matched them to disciplinary classification by journal, the Scopus ASJC classification. To focus on relatively high signal, we only considered Scopus ASJC minor subject area categories that accounted for at least 15\% of the publications in each cluster.

@@

We also attempted to map these 20 clusters assembled through direct citation to the CCS classification by manually matching the top 25-50 most heavily cited publications in each cluster to corresponding categories in the CCS. This was feasible in some cases, clusters 0 and 1 mapped reasonably well to the top level categories Hardware and Computer Systems Organization but less so in the cases of other clusters. Further, the top cited papers often derived from biology and we were concerned that this small sample (top cited papers) were not representative of the cluster. [Overall conclusions of Graclus based clustering?]

\subsection{Clustering by Co-citation.} [Sentence justifying co-citation base clustering]

@@

[Overall conclusions of cocitation based clustering]

@@

\begin{figure}[ht]
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.45]{scopus_dblp_heatmap3.pdf}
% figure caption is below the figure
\caption{Heat map for the clustering obtained by direct citation. The  y-axis (rows) correspond to topics, defined by Scopus characterizations, and the
x-axis (columns) represent the 20 different clusters.  
Each cluster is characterized by topics that label at least 10\% of the publications in the cluster.
 }
\label{fig:heatmap}       % Give a unique label
\end{figure}
\newpage

\begin{figure}[ht]
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.45]{graclus_cocit_fig.pdf}
% figure caption is below the figure
\caption{Heat map for the clustering obtained by direct citation. The  y-axis (rows) correspond to topics, defined by Scopus characterizations, and the
x-axis (columns) represent the 20 different clusters.  
Each cluster is characterized by topics that label at least 10\% of the publications in the cluster.
 }
\label{fig:heatmap}       % Give a unique label
\end{figure}

@@

\begin{table}[ht]
\caption{\emph{Statistics regarding the 20 clusters produced using direct citation. The \emph{comp} dataset 
2685021 is because of inner joins on ASJC Explain conductance calculations~\cite{shun_parallel_2016}}}
\label{tab:1}       
\begin{tabular}{lrccc}
\hline\noalign{\smallskip}
minor\_subject\_area & Percent of Publications \\
\noalign{\smallskip}\hline\noalign{\smallskip}
Software & 30.30 \\ 
Computer Science Applications & 25.20 \\ 
Computer Networks and Communications & 21.60 \\ 
Computer Science(all) & 19.60 \\ 
Artificial Intelligence & 14.30 \\ 
Information Systems & 11.90 \\ 
Hardware and Architecture & 11.50 \\ 
Signal Processing & 9.90 \\ 
Computer Vision and Pattern Recognition & 9.50 \\ 
Computational Theory and Mathematics & 8.00 \\ 
Human-Computer Interaction & 7.50 \\ 
Computer Graphics and Computer-Aided Design & 5.20 \\ 
Computer Science (miscellaneous) & 0.90 \\ 
\noalign{\smallskip}\hline
\end{tabular}
\end{table}
\newpage
